Let's Get Movin': Frontiers of Optimal Transport

### 2018-04-26 reports/paper/paper.tex
.
Add MNIST section to paper

We may then benchmark the performance of this data by generating data. We
generate random square $n \times n$ cost matrices by making each entry i.i.d.
$\chi^2(1)$, and perform optimal transport on $\mathbf{a} = \mathbf{b} =
\mathbf{1}_n / n$. The runtimes for different values of $n$ are documented in
\cref{fig:lpruntime}.

\begin{figure}[h]
\includegraphics[width=\textwidth]{../../code/lpruntime}
\caption{Runtime comparison}
\label{fig:lpruntime}
\end{figure}

@@

@@

>>>>

\begin{figure}[b]
\includegraphics[width=\textwidth]{../../code/lpruntime}
\caption{LP Solver Runtime}
\label{fig:lpruntime}
\end{figure}

We may then benchmark the performance of this data by generating data. We
generate random square $n \times n$ cost matrices by making each entry i.i.d.
$\chi^2(1)$, and perform optimal transport on $\mathbf{a} = \mathbf{b} =
\mathbf{1}_n / n$. The runtimes for different values of $n$ are documented in
\cref{fig:lpruntime}. While this gives us a baseline to how quickly we can
compute optimal transport problems, solving the problem for $n$ bigger than a
couple dozen quickly becomes an expensive task. Thus, we may look elsewhere for
more efficient solutions that either solve a special case or provide an
approximate solution.

@@

One such special case of the optimal transport problem is when $\mathbf{a} =
\mathbf{b} \propto \mathbf{1}_n$ as in our simulations to test the performance
of the generic LP solver. In this case, mass is being matched in unit pieces,
which makes the problem more tractable. The Hungarian algorithm solves this
special-case.

% TODO: exposition on how Hungarian works

\begin{figure}[h]
\includegraphics[width=\textwidth]{../../code/hungarianruntime.pdf}
\caption{Hungariant Algorithm runtime}
\label{fig:hungarianruntime}
\end{figure}

@@

The applications of the optimal transport problem to the physical transportation
of objects is practically explicit in the problem statement. This is but one of
many applications, though. In particular, we explore the use of optimal
transport in the image processing of MNIST images.

\begin{figure}[tb]
  \centering
  \includegraphics{../../code/ot-mnist-evol.pdf}
  \caption{Optimal transport on MNIST images}
  \label{fig:mnisttransport}
\end{figure}

Optimal transport provides a natural language with which to talk about image
similarity. If we consider the darkness of a pixel as ``mass,''  the similarity
of two images could be quantified as the units of mass times distance that need
to be moved in order to transform it into the second image. This requires us to
normalize the image such that the pixel intensities sum to 1, but this is not a
problem with MNIST images which do not differ significantly in their average
brightness. The transport also requires a choice of cost function between two
pixels. We opt for an $L^2$ penalization on the distance between the pixels, but
alternative metrics such as $L^1$ perform similarly \cref{fig:mnisttransport}
illustrates what this transformation looks like for a ``0'' being transported
into a ``6.''

To verify that the similarity is behaving as we expect, we can compare the
similarity of a number ``5'' against its most similar images in a small sample
of the MNIST dataset. The most similar images are also 5's, as depicted in
\cref{fig:mnistsim}.

\begin{figure}[tb]
  \centering
  \includegraphics{../../code/mnist-similarity.pdf}
  \caption{MNIST Similarity Metric}
  \label{fig:mnistsim}
\end{figure}
